{\fontsize{14.7}{16}\selectfont Score-Based Tests of Measurement Invariance with Respect to Continuous and Ordinal Variables}

### 2015-07-23 ecpa2015/slides.Rnw
small touch-ups
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@378 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\textbf{Crucial assumption:} Measurement invariance (MI). Otherwise observed differences in
scores cannot be reliably attributed to the latent variable that the model
purports to measure.

@@

\textbf{Problems:}
\begin{itemize}
  \item Subgroups have to be formed in advance.
  \item Numerical variables are often categorized into groups in an ad hoc way
    (e.g., splitting at the median).
  \item In ordinal variables the ordering of the categories is often not exploited
    (assessing only if at least one group differs from the others).
  \item When likelihood ratio or Wald tests are employed, the model has to be
    fitted to each subgroup which can become numerically challenging and computationally
    intensive.
\end{itemize}

@@

\textbf{Idea:}
\begin{itemize}
  \item Generalize the LM test (or score test).
  \item Thus, the model only has to be fitted once under the MI assumption to the full data set.
  \item Catpure model deviations along a variable that is suspected to cause MI violations.
  \item Exploit ordering to assess if there is (at least) one split so that the
    model parameters before and after the split differ.
  \item The split does \emph{not} have to be known or guessed in advance.
\end{itemize}

@@

\textbf{Illustration:} CFA for artificial data.
\begin{itemize}
  \item Model with two latent scales: Verbal and math score.
  \item Three manifest variables for each scale.
  \item Violation of MI for the math loadings along the age of the subjects.
\end{itemize}

@@

\begin{itemize}
  \item General score-based test framework for assessing measurement
    invariance in parametric psychometric models.
  \item Assessment is along some variable $v$ which can be continuous,
    ordinal, or categorical.
  \item Tests can be seen as generalizations of the Lagrange multiplier
    (or score) test.
  \item Computation of critical values might require simulation from
    certain stochastic processes (Brownian bridges).
  \item Easy-to-use implementation available in R package \emph{strucchange}.
  \item Can be re-used in model-based recursive partitioning.
\end{itemize}

>>>>

\textbf{Crucial assumption:} Measurement invariance (MI). Otherwise observed differences in
scales cannot be reliably attributed to the latent variable that the model
purports to measure.

@@

\textbf{Problems:}
\begin{itemize}
  \item Subgroups have to be formed in advance.
  \item Continuous variables are often categorized into groups in an ad hoc way
    (e.g., splitting at the median).
  \item In ordinal variables the ordering of the categories is often not exploited
    (assessing only if at least one group differs from the others).
  \item When likelihood ratio or Wald tests are employed, the model has to be
    fitted to each subgroup which can become numerically challenging and computationally
    intensive.
\end{itemize}

@@

\textbf{Idea:}
\begin{itemize}
  \item Generalize the LM test.
  \item Thus, the model only has to be fitted once under the MI assumption to the full data set.
  \item Catpure model deviations along a variable that is suspected to cause MI violations.
  \item Exploit ordering to assess if there is (at least) one split so that the
    model parameters before and after the split differ.
  \item The split does \emph{not} have to be known or guessed in advance.
\end{itemize}

@@

\textbf{Illustration:} CFA for artificial data.
\begin{itemize}
  \item Model with two latent scales (verbal and math).
  \item Three manifest variables for each scale.
  \item Violation of MI for the math loadings along the age of the subjects.
\end{itemize}

@@

\begin{itemize}
  \item General score-based test framework for assessing measurement
    invariance in parametric psychometric models.
  \item Assessment is along some variable $v$ which can be continuous,
    ordinal, or categorical.
  \item Tests can be seen as generalizations of the Lagrange multiplier
    test.
  \item Computation of critical values might require simulation from
    certain stochastic processes (Brownian bridges).
  \item Easy-to-use implementation available in R package \emph{strucchange}.
  \item Can be re-used in model-based recursive partitioning.
\end{itemize}
